---
title: Process monthly precipitation climatologies from the GHCN monthly database
author: Adam M. Wilson
output:
  knitrBootstrap::bootstrap_document:
    title: "GHCN Compare"
    theme: amelia
    highlight: sunburst
    theme.chooser: TRUE
    highlight.chooser: TRUE
---    


```{r setup,echo=F,message=F}
source("/media/data/Cloud/analysis/setup.R")
opts_knit$set(root.dir=getwd(),cache=T,base.url = NULL,fig.width=8)
```

# GHCN Data Processing
## Download, uncompress, and import GHCN data
```{r datadownload}
if(!file.exists("data/ghcn/v2.prcp")){
system("wget ftp://ftp.ncdc.noaa.gov/pub/data/ghcn/v2/v2.prcp.Z -P data/ghcn/")
system("uncompress data/ghcn/v2.prcp.Z")
}

if(!file.exists("data/ghcn/v2.prcp.inv"))
   system("wget ftp://ftp.ncdc.noaa.gov/pub/data/ghcn/v2/v2.prcp.inv -P data/ghcn/")

```


## GHCN pre-processing

Subset to 2000-present, take monthly means, & reshape to long format

```{r dataprocess}
fpds="data/out/ghcn.csv"
if(!file.exists(fpds)){

## read in metadata
fwi=c(11,-1,20,10,8,7,5)
pdi=read.fwf("data/ghcn/v2.prcp.inv",
            widths=fwi,col.names=c("id","stationname","country","latitude","longitude","elevation"),comment.char="")

## read in precipitation data
fwd=c(11,1,4,rep(5,12))
pd=read.fwf("data/ghcn/v2.prcp",na.strings=-9999,
             widths=fwd,col.names=c("id","duplicate","year",month.name),comment.char="")


pdc1=pd%.%filter(year>=1950);pdc1$era="1950-2014"
pdc2=pd%.%filter(year>=2000);pdc2$era="2000-2014"

pdl=melt(rbind(pdc1,pdc2),id.vars=c("id","year","era"),measure.vars=month.name)

#convert "trace" precipitation to 0.1mm
pdl$value[pdl$value==-8888]=1

pds=pdl%.%group_by(id,era,variable)%.%summarize(
  mmp=mean(value/10,na.rm=T),
  count=sum(!is.na(value)),
  min=min(value/10,na.rm=T),
  max=max(value/10,na.rm=T))

pds$month=1:12[pds$variable==month.name]
pds[,colnames(pdi)[-1]]=pdi[match(pds$id,pdi$id),-1]
pds=as.data.frame(pds)

## drop stations not in inventory data
pds=pds[!is.na(pds$latitude),]

## extract cloud data
cf=stack(list.files("data/MCD09/",pattern="MCD09_mean_",full=T))
NAvalue(cf)=0
gain(cf)=.01

pts=unique(pds[,c("id","latitude","longitude")])
coordinates(pts)=c("longitude","latitude")

biome=raster("data/out/teow.tif")
pts=raster::extract(biome,pts,sp=T)

pts=raster::extract(cf,pts,sp=T)
ptsl=melt(pts@data,id.var=c("id","teow"))
ptsl$month=as.numeric(do.call(rbind,strsplit(as.character(ptsl$variable),"_"))[,3])
pds$cld=ptsl$value[match(paste(pds$id,pds$month),paste(ptsl$id,ptsl$month))]
pds$biome=pts$teow[match(pds$id,pts$id)]

write.csv(pds,file=fpds,row.names=F)
}
pds=read.csv(fpds)
pds$elevation[pds$elevation==-999]=NA
pds$variable=factor(pds$variable,levels=month.name,ordered=T)
```

## Comparison with Precipitation 

Mean monthly precipitation over 1950-2014
```{r mmpmap,fig.width=15}
ggplot(pds,aes(x=longitude,y=latitude,colour=mmp+1,order = mmp))+
  geom_point(size=1)+
  facet_wrap(~variable)+
  scale_colour_gradientn(trans = "log",colours=c("grey","tan","darkgreen","green","purple"))+
  coord_fixed(ratio = 1.25)
```

## Mean monthly cloud frequency and precipitation

```{r cloudAndPrecip,fig.width=10,fig.height=15}
ggplot(na.omit(pds),aes(x=cld,y=mmp+1))+
  stat_density2d(geom="tile", aes(fill = ..density..), contour = FALSE)+
  facet_wrap(~variable)+
  scale_fill_gradientn(colours=c("transparent","blue","red","purple"))+
  stat_smooth(method="loess",se=T,col="black")+
  scale_y_continuous(trans = log10_trans(),
                        breaks = trans_breaks("log10", function(x) 10^x),
                        labels = trans_format("log10", math_format(10^.x)))+
  ylab("Mean Monthly Precipitation (mm)")+
  xlab("Mean Monthly Cloud Frequency (%)")
  
```

## Mean monthly precipitation vs. cloud frequency by biome

Global Spearman's Correlation between mean monthly temperature and precipitation:
```{r globalcor,results='asis'}
t1=cor.test(pds$mmp[pds$era=="1950-2014"] , pds$cld[pds$era=="1950-2014"] , method = "spearman",use="complete")
t1n=nrow(na.omit(data.frame(pds$mmp[pds$era=="1950-2014"] , pds$cld[pds$era=="1950-2014"])))
t2=cor.test(pds$mmp[pds$era=="2000-2014"] , pds$cld[pds$era=="2000-2014"] , method = "spearman",use="complete")
t2n=nrow(na.omit(data.frame(pds$mmp[pds$era=="2000-2014"] , pds$cld[pds$era=="2000-2014"])))

td=data.frame(Era=c("1950-2014","2000-2014"),
              "SpearmanRho"=c(t1$estimate,t2$estimate),
              'p-value'=c(t1$p.value,t2$p.value),
              n=c(t1n,t2n))
kable(td)
```

```{r summarizebybiome,fig.width=10,fig.height=15}
tb=table(pds$biome)
pds$biome=as.factor(pds$biome)
bcode=read.csv("data/out/biomecode_condensed.csv")

pds$biomename=paste(bcode$realm,bcode$biome,sep="_")[match(pds$biome,bcode$icode)]

pdss=data.frame(cor=c(by(data=pds,INDICES=list(pds$biomename),FUN=function(x){
  tc=cor(x$cld,log(x$mmp+1),method="spearman",use="complete")
  round(tc,2)
},simplify=T)))
pdss$biomename=rownames(pdss);rownames(pdss)=1:nrow(pdss)

ggplot(na.omit(pds[pds$biome%in%names(tb)[tb>100],]),aes(x=cld,y=mmp+1))+
  facet_wrap(~biomename)+
  stat_density2d(geom="tile", aes(fill = ..density..), contour = FALSE)+
  scale_fill_gradientn(colours=c("transparent","blue","red","purple"))+
  stat_smooth(method="loess",se=T,col="black")+
  scale_y_continuous(trans = log10_trans(),
                        breaks = trans_breaks("log10", function(x) 10^x),
                        labels = trans_format("log10", math_format(10^.x)))+
  ylab("Mean Monthly Precipitation (mm)")+
  xlab("Mean Monthly Cloud Frequency (%)")+
  geom_text(data = pdss, aes(x = 20, y = 1000, label = cor, family = "serif"), 
          color = 'blue',  parse = TRUE)

```

```{r modeleval,results='asis'}
m1=lm(cld~mmp,data=pds)
m2=lm(cld~elevation,data=pds)
m3=lm(cld~log(mmp+1)+elevation,data=pds)
m4=lm(cld~log(mmp+1)+elevation,data=pds)
m5=lm(cld~log(mmp+1)+biomename,data=pds)

#
screenreg(list(m1,m2,m3,m4,m5),ci.force=T,single.row=T,omit.coef="biome")
htmlreg(list(m1,m2,m3,m4,m5),ci.force=T,single.row=T,doctype=F,omit.coef="biome")
```